Water, High-Altitude Condensates, and Possible Methane Depletion in the Atmosphere of the Warm Super-Neptune WASP-107\lowercase{b}

### 2018-04-30 arxiv_v4_accepted/ms.tex
accepted version with proof corrections

The best fit light curves have a median $\chi^2_\nu$ value of 1.16.  We conservatively rescaled the photometric uncertainties for all spectroscopic channels such that the $\chi^2_\nu$ values are unity. We performed an MCMC fit to the light curves with \texttt{emcee}.  For each light curve we ran a fit with 50 walkers and 1000 steps per walker, and tested for convergence as we did for the white light curve. The median transit depths and $1\,\sigma$ uncertainties are given in Table\,\ref{tab:tspec}.

@@

\begin{figure*}
\includegraphics[width = \textwidth]{fig2_spectrum.pdf}
\caption{The transmission spectrum of WASP-107b (points with 1\,$\sigma$ error bars) compared to retrieved models (blue line with shaded $1$ and $2\sigma$ confidence intervals). The features at $1.15$ and $1.4$ $\mu$m are due to water absorption. The right-hand axis indicates the normalized atmospheric scale height assuming solar composition. The water feature crosses less than one scale height, indicating that condensates are present at high altitude. \added{We also show the best fit model from an analysis where the C/O is fixed at 0.54, the solar value (dashed line). There are subtle differences in feature shape due to methane absorption in the solar C/O case, which cause the retrieval to favor sub-solar C/O values at 2.7$\sigma$ confidence.}} 
\label{fig:spectrum}
\end{figure*}

@@

The nominal retrieval results are shown in Figure\,\ref{fig:retrieval}.  The best fit model provides a good fit to the data ($\chi^2_\nu = 1.2$).  The metallicity distribution spans the full range allowed by our priors, with preference for larger values. The cloud top pressure is estimated to be $0.01 - 3$ mbar at $1\,\sigma$ confidence. The retrieved irradiation temperature is $525 - 820$ K (1 $\sigma$ confidence).  We find that the C/O value is less than solar (0.54) at $2.7\,\sigma$ confidence.  We tested the detection significance for water by removing water opacity from the nominal model. The Bayesian evidence favors the inclusion of water at $6.5\,\sigma$ confidence.

We explored a few retrieval scenarios with additional complexity, including the addition of cloud patchiness \citep{line16}, a quench pressure for nitrogen and carbon species \citep[e.g.][]{morley17}, a power law haze opacity, \added{no prior on atmospheric metallicity, a more flexible T/P profile (with added free parameters for infrared opacity and visible-to-infrared opacity), and the more flexible T/P profile with quenching}.  We also varied the assumed planet mass by $3\,\sigma$. These models did not significantly improve the fit quality, and the \added{retrieved cloud-top pressure and chemical composition were consistent to within $0.1\,\sigma$}.

\added{For example, when we removed the $30\times$ solar metallicity prior upper limit,} the retrieved metallicity is $[M/H] = 1.35^{+0.68}_{-1.48}$, log10(C/O) is $-1.57^{+0.54}_{-0.31}$, and cloud top pressure $\log P_\mathrm{c} = -3.63^{+0.81}_{-0.48}.$ These results show that the inferred C/O ratio and cloud-top pressure are unaffected by the prior on atmospheric metallicity.  The metallicity is consistent with the interior structure modeling (which puts a $3\,\sigma$ upper limit on metallicity of $[\mathrm{M/H}] = 1.5$). The transmission spectrum does not constrain the atmospheric metallicity more precisely because there is a degeneracy between metallicity and the planet's 10-bar radius \citep[e.g]{griffith14,heng17} with the constraint primarily driven by the H$_2$O-H$_2$ CIA continuum ratio, molecular weight \citep{line16}, and pressure broadening \citep{dewit13}.

\added{Similarly, we find that the transmission spectrum does not provide a strong constraint on the shape of the temperature pressure profile. The primary constraint on temperature structure is the effective scale height temperature, which sets the amplitude of features in the spectrum. To confirm this, we explored using a more flexible T/P profile with free parameters for infrared and visible-to-infrared opacity, in addition to irradiation temperature (which we label case 2). Compared to our nominal model (case 1), which just varies irradiation temperature but keeps the T/P profile shape fixed, we found that the retrieved temperature at the photosphere is nearly identical. For example, at $P = 10^{-4}$ bar (the most probable cloud-top pressure), we find $T = 618^{+150}_{-120}$ K for case 1, and $T = 615^{+178}_{-123}$ K for case 2. The constraints on chemical composition agree at the 0.1-$\sigma$ level for the two cases.}

@@

\acknowledgments
We thank Fei Dai, Jessica Spake, Ian Crossfield, Hannah Diamond-Lowe, and Jonathan Fortney for productive conversations. L.K. acknowledges support from the Harvard Society of Fellows and the Harvard Astronomy Department Institute for Theory and Computation. C.V.M. acknowledges support from NASA through the Sagan Fellowship Program.

@@

\bibitem[{{Gao} {et~al.}(2017){Gao}, {Marley}, {Zahnle}, {Robinson}, \&
  {Lewis}}]{gao17}
{Gao}, P., {Marley}, M.~S., {Zahnle}, K., {Robinson}, T.~D., \& {Lewis}, N.~K.
  2017, \aj, 153, 139

@@

\bibitem[{{Zahnle} {et~al.}(2009){Zahnle}, {Marley}, {Freedman}, {Lodders}, \&
  {Fortney}}]{zahnle09}
{Zahnle}, K., {Marley}, M.~S., {Freedman}, R.~S., {Lodders}, K., \& {Fortney},
  J.~J. 2009, \apjl, 701, L20

>>>>

The best fit light curves have a median $\chi^2_\nu$ value of 1.16.  We
conservatively rescaled the photometric uncertainties for all spectroscopic
channels such that the $\chi^2_\nu$ values are unity. We performed an MCMC fit
to the light curves with \texttt{emcee}.  For each light curve we ran a fit with
50 walkers and 1000 steps per walker, and tested for convergence as we did for
the white light curve. The median transit depths and $1\,\sigma$ uncertainties
are given in Table\,\ref{tab:tspec}, and the transmission spectrum is shown in
Figure\,\ref{fig:spectrum}.

@@

\begin{figure*}
\includegraphics[width = \textwidth]{fig2_spectrum.pdf}
\caption{The transmission spectrum of WASP-107b (points with 1\,$\sigma$ error bars) compared to retrieved models (blue line with shaded $1$ and $2\sigma$ confidence intervals). The features at $1.15$ and $1.4$ $\mu$m are due to water absorption. The right-hand axis indicates the normalized atmospheric scale height assuming solar composition. The water feature crosses less than one scale height, indicating that condensates are present at high altitude. We also show the best fit model from an analysis where the C/O is fixed at 0.54, the solar value (dashed line). There are subtle differences in feature shape due to methane absorption in the solar C/O case, which cause the retrieval to favor sub-solar C/O values at 2.7$\sigma$ confidence.} 
\label{fig:spectrum}
\end{figure*}

@@

The best-fit spectrum is shown in Figure\,\ref{fig:spectrum}, and the nominal retrieval results are shown in Figure\,\ref{fig:retrieval}.  The best fit model provides a good fit to the data ($\chi^2_\nu = 1.2$).  The metallicity distribution spans the full range allowed by our priors, with preference for larger values. The cloud top pressure is estimated to be $0.01 - 3$ mbar at $1\,\sigma$ confidence. The retrieved irradiation temperature is $525 - 820$ K (1 $\sigma$ confidence).  We find that the C/O value is less than solar (0.54) at $2.7\,\sigma$ confidence.  We tested the detection significance for water by removing water opacity from the nominal model. The Bayesian evidence favors the inclusion of water at $6.5\,\sigma$ confidence.

We explored a few retrieval scenarios with additional complexity, including the addition of cloud patchiness \citep{line16}, a quench pressure for nitrogen and carbon species \citep[e.g.][]{morley17}, a power law haze opacity, no prior on atmospheric metallicity, a more flexible T/P profile (with added free parameters for infrared opacity and visible-to-infrared opacity), and the more flexible T/P profile with quenching.  We also varied the assumed planet mass by $3\,\sigma$. These models did not significantly improve the fit quality, and the retrieved cloud-top pressure and chemical composition were consistent to within $0.1\,\sigma$.

For example, when we removed the $30\times$ solar metallicity prior upper limit, the retrieved metallicity is $[M/H] = 1.35^{+0.68}_{-1.48}$, log10(C/O) is $-1.57^{+0.54}_{-0.31}$, and cloud top pressure $\log P_\mathrm{c} = -3.63^{+0.81}_{-0.48}.$ These results show that the inferred C/O ratio and cloud-top pressure are unaffected by the prior on atmospheric metallicity.  The metallicity is consistent with the interior structure modeling (which puts a $3\,\sigma$ upper limit on metallicity of $[\mathrm{M/H}] = 1.5$). The transmission spectrum does not constrain the atmospheric metallicity more precisely because there is a degeneracy between metallicity and the planet's 10-bar radius \citep[e.g]{griffith14,heng17} with the constraint primarily driven by the H$_2$O-H$_2$ CIA continuum ratio, molecular weight \citep{line16}, and pressure broadening \citep{dewit13}.

Similarly, we find that the transmission spectrum does not provide a strong constraint on the shape of the temperature pressure profile. The primary constraint on temperature structure is the effective scale height temperature, which sets the amplitude of features in the spectrum. To confirm this, we explored using a more flexible T/P profile with free parameters for infrared and visible-to-infrared opacity, in addition to irradiation temperature (which we label case 2). Compared to our nominal model (case 1), which just varies irradiation temperature but keeps the T/P profile shape fixed, we found that the retrieved temperature at the photosphere is nearly identical. For example, at $P = 10^{-4}$ bar (the most probable cloud-top pressure), we find $T = 618^{+150}_{-120}$ K for case 1, and $T = 615^{+178}_{-123}$ K for case 2. The constraints on chemical composition agree at the 0.1-$\sigma$ level for the two cases.

@@

\acknowledgments
We thank Fei Dai, Jessica Spake, Ian Crossfield, Hannah Diamond-Lowe, and Jonathan Fortney for productive conversations. L.K. acknowledges support from the Harvard Society of Fellows and the Harvard Astronomy Department Institute for Theory and Computation. C.V.M. acknowledges support from NASA through the Sagan Fellowship Program. M.R.L. acknowledges NASA XRP grant NNX17AB56G for partial support of the theoretical interpretation of the data, as well as the ASU Research Computing staff for support with the Saguaro and Agave compute clusters.

@@

@@
